\documentclass[10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[english,russian]{babel}
\usepackage[left=10mm, right=10mm, top=10mm, bottom=10mm]{geometry}

%Show bad boxes
\overfullrule=5pt

% DO indent first paragraph
\usepackage{indentfirst}

% Two-columns
\usepackage{multicol}

% Cells spanning several row in tables
\usepackage{multirow}


\usepackage[normalem]{ulem}

% Disable page numbering
\pagenumbering{gobble}

% Do not indent lists and for
\usepackage{enumitem}
\setlist{nosep}

% ==========================================
% Store original par indent for future uses
% ==========================================
\newlength{\OriginalParIndent}
\OriginalParIndent=\the\parindent % Remember original parindent
%\showthe\parindent
%\showthe\tmp
% ==========================================

% Paragraph tweaks
\setlength{\parindent}{0pt} % Do not indent every paragraph
\setlength{\parskip}{1.1mm}   % More space between paragraphs

% Less space between lines
\renewcommand{\baselinestretch}{0.97}

% Domain specific commands
\newcommand\ex[1]{\textit{\textbf{{#1}}}}           % Example
\newcommand\sex[1]{\textit{\textbf{{\small {#1}}}}} % Small Example
\newcommand\za{\_\thinspace }                       % Zero Article


% Itemize with less indentation
\newenvironment{ItemizeWithOrigParIndent}
    {\begin{itemize}[leftmargin=\OriginalParIndent]}
    {\end{itemize}}


%\newenvironment{NextLineDescription}[1]
%    {\begin{description}[#1,style=nextline,before={\renewcommand\makelabel[1]{##1~---}}]}
%    {\end{description}}


% ==========================================
% Show two column boundaries
% ==========================================
%\usepackage{etoolbox}
%\newlength\Fcolumnseprule
%\setlength\Fcolumnseprule{0.4pt}
%\makeatletter
%\newcommand\ShowInterColumnFrame{
%\patchcmd{\@outputdblcol}
%  {{\normalcolor\vrule \@width\columnseprule}}
%  {\vrule \@width\Fcolumnseprule\hfil
%    {\normalcolor\vrule \@width\columnseprule}
%    \hfil\vrule \@width\Fcolumnseprule
%  }
%  {}
%  {}
%}
%\makeatother
%\ShowInterColumnFrame
% ==========================================



\hyphenation{band-wa-gon}
\hyphenation{Ro-yal}

\begin{document}

% Thinner bullet in lists see
% http://tex.stackexchange.com/questions/8510/reduce-size-of-bullet-character-in-lists
\renewcommand{\labelitemi}{$\vcenter{\hbox{\tiny$\bullet$}}$}


%================================================================================================
%================================================================================================
%================================================================================================
\setcounter{secnumdepth}{1} % Enumerate only Sections (1)
\begin{multicols}{2}[\section{SpeakOut Intermediate}]

\subsection{Unit 1}

\subsubsection{Vocabulary: Idioms}
\begin{description}[leftmargin=5cm,style=nextline,before={\renewcommand\makelabel[1]{##1~---}}]
  \item[\ex{Take part in something}] принять участие в чём-то
  \item[\ex{Take after someone}] родиться похожим на
  \item[\ex{Take responsability for}] брать ответсвенность за
  \item[\ex{Get on with someone}] ладить с
  \item[\ex{Do someone a favour}] сделать одолжение
  \item[\ex{Ask a favour of smb}] ask a favour of my parents
\end{description}



%================================================================================================
%================================================================================================
%================================================================================================
\subsection{Unit 2}
\subsubsection{Preposition + expressions of time}
To help you remember which preposition of time to use, try to memorise this: \textbf{on} M\textbf{on}day; \textbf{in} w\textbf{in}ter; \textbf{at} th\textbf{at} time.

\begin{description}[leftmargin=1.2cm,style=nextline,before={\renewcommand\makelabel[1]{##1~=}}]
  \item[\textbf{on}] for specific days,
  \item[\textbf{in}] for time periods,
  \item[\textbf{at}] for specific times.
\end{description}

\begin{description}[leftmargin=2.8cm,style=nextline,before={\renewcommand\makelabel[1]{##1~---}}]
  \item[\ex{On my own}] alone, not whith other people.
  \item[\ex{On purpose}] it was not mistake.
  \item[\ex{In a hurry}] cannot wait.
\end{description}


\subsubsection{Vocabulary}
\begin{description}[leftmargin=2.3cm,style=nextline,before={\renewcommand\makelabel[1]{##1~---}}]
  \item[\ex{Violent}] насильственный.
  \item[\ex{Strike}] забастовка, удар.
  \item[\ex{Fugitive}] беглец.
  \item[\ex{Hostages}] заложники.
  \item[\ex{Prisoner}] заключённый.
\end{description}



%================================================================================================
%================================================================================================
%================================================================================================
\subsection{Unit 3}
\subsubsection{Vocabulary: Idioms}
\begin{description}[leftmargin=4.5cm,style=nextline,before={\renewcommand\makelabel[1]{##1~---}}]
  \item[\ex{We are running out of time}] время на исходе.
\end{description}
\vspace{-\parskip}
\begin{description}[leftmargin=4.7cm,style=nextline,before={\renewcommand\makelabel[1]{##1~---}}]
  \item[\ex{Go window shopping}] разглядывать ветрины, прогуливаться по магазинам.
  \item[\ex{Let your hair down}] позволить себе вести себя более спокойно и расслаблено.
  \item[\ex{Be in two minds}] колебаться, не знать на что решиться.
  \item[\ex{Work against the clock}] работать не покладая рук.
  \item[\ex{It's not my cup of tea}] это не в моём вкусе.
  \item[\ex{Close to my heart}] близко моему сердцу.
  \item[\ex{Put my foot in it}] дать маху.
\end{description}
\vspace{-\parskip}
\begin{description}[leftmargin=3.3cm,style=nextline,before={\renewcommand\makelabel[1]{##1~---}}]
  \item[\ex{Break the ice}] положить начало.
  \item[\ex{Learn by heart}] учить на изусть.
  \item[\ex{Travel light}] путешествовать на-легке.
  \item[\ex{Go grey}] седеть.
  \item[\ex{Small talk}] болтовня.
  \item[\ex{On our mind}] на уме.
  \item[\ex{In hot water}] в беде (по своей вине).
\end{description}



%================================================================================================
%================================================================================================
%================================================================================================
\subsection{Unit 4}
\subsubsection{Vocabulary: Confusing words}

\begingroup
\parindent=\OriginalParIndent

\ex{job~--- work}. \textbf{Work} is what you do to earn money: \textit{What kind of \ex{work} does he do?}
A \textbf{job} is the particular type of work that you do: \textit{Sam's got a \ex{job} as a waiter}.
\textbf{Job} can be plural, but \textbf{work} cannot.

\ex{remember~--- remind}. If you \textbf{remember} something, a fact or event from the past,
or something you earlier decided to do, comes back into you mind:
\textit{He suddenly \ex{remembered} he had to go to the bank}. If someone or something \textbf{reminds} you
of something, they make you remember it:
\textit{\ex{Remind} me to call him later today}.


\ex{forget~--- leave}. If you want to talk about the place where you have left something, use the
verb \textbf{leave}, not the verb \textbf{forget}. Compare: \textit{I've \ex{forgotten} my book and I've
\ex{forgotten} my keys.} \textit{I've \ex{left} my keys in the car}.\\
Don't say: \textit{\sout{I've forgotten my keys in the car.}}

\ex{hear~--- listen}.

\ex{earn~--- win}.

\ex{fun~--- funny}. Use \textbf{fun} to talk about events and activities that are enjoyable, such as
games and parties: \textit{Let's go to the beach and have some \ex{fun}.} \textbf{Funny} is an adjective
that describes someone or something that makes you laugh: \textit{Bob's jokes are really \ex{funny}.}

\ex{miss~--- lose}. \textbf{miss}~--- пропустить; \textbf{lose}~--- потерять.

\endgroup


\subsubsection{Vocabulary: VERY --> without VERY}
\begingroup
\setlength{\columnsep}{19pt}
\begin{multicols}{2}
\begin{description}[leftmargin=1.6cm,style=nextline,before={\renewcommand\makelabel[1]{##1~---}}]
  \item[good] wonderful, amazing, brilliant
  \item[big] enormous, giant
  \item[small] tiny
  \item[tired] exhausted
  \item[hot] boiling
  \item[cold] freezing
  \item[tasty] delicious
  \item[angry] furious
  \item[pretty] beautyful
\end{description}
\vspace{-\parskip}
\begin{description}[leftmargin=2.3cm,style=nextline,before={\renewcommand\makelabel[1]{##1~---}}]
  \item[interesting] fascinating
  \item[difficult] impossible
\end{description}
\end{multicols}
\endgroup



\subsubsection{Vocabulary: Covering letter}
\begin{ItemizeWithOrigParIndent}
  \item \ex{Regarding your advertisement\dots}~--- about
  \item \ex{I would like to submit an application for the\dots}~--- to apply for the job
  \item \ex{I meet all the requirements\dots}~--- I think I would be good for the job
  \item \ex{Proven ability in\dots}~--- I have shown that I am able to do this
  \item \ex{I look forward to hearing from you at your earliest convenience}~--- as soon as you
  have the opportunity
\end{ItemizeWithOrigParIndent}






%================================================================================================
%================================================================================================
%================================================================================================
\subsection{Unit 7}
\subsubsection{Vocabulary: Success 1}
\begin{ItemizeWithOrigParIndent}
    \item work hard \ex{at}
    \item focus \ex{on}
    \item get better \ex{at}
    \item have the opportunity \ex{to} do something
    \item practise somehting
\end{ItemizeWithOrigParIndent}

\subsubsection{Vocabulary: Success 2}
\begin{ItemizeWithOrigParIndent}
    \item Bloggers \ex{pay attention to} what is going on in the world.
    \item We \ex{pay for} the pizza.
    \item He is \ex{waiting for} her gilgriend.
    \item She \ex{has a talent for} finding a bargain.
    \item He has an \ex{aptitude for} maths.
    \item The film \ex{picks up on} the difficulties people experienced during the war.
    \item He works very hard. He \ex{puts in} a lot of hours.
    \item I am \ex{usless at} ball sport.
    \item I am \ex{hopeless at} geography.
    \item He is \ex{gifted at} painting.
    \item He is an \ex{expert in} Italian art.
    \item He has a lot of \ex{ability as} a dancer.
\end{ItemizeWithOrigParIndent}

\subsubsection{Vocabulary: Clarifying Opinions}
\begin{ItemizeWithOrigParIndent}
    \item For one thing, \dots\ For another, \dots
    \item For me \sout{To me}
    \item Having said that
    \item Like I said
\end{ItemizeWithOrigParIndent}





%================================================================================================
%================================================================================================
%================================================================================================
\subsection{Unit 8}
\subsubsection{Grammar: Relative Clause}
There are two types of relative clauses: defining and non-defining ones.

\textbf{Defining relative clauses:}
\begin{ItemizeWithOrigParIndent}
   \item We can leave out \textit{who}, \textit{which} and \textit{what} when these words are
   not the subject of the relative clause: \textit{She is the girl (who) I saw yesterday}.
   \item Do not use commas around it.
   \item Spell it fast, without pause in speech.
\end{ItemizeWithOrigParIndent}

They act like a filter (or a function) which, given an input set, returns a smaller / narrower subset,
ideally consisting of one item.

For example: \textit{The town where I live is full of trees}. The input set here is all towns
of the world, the function filters out all towns except for the town where he lives in,
and returns a smaller subset consisting of one town where he currently lives.

You can see how input set was reduced to a smaller subset out of just one item.

\textbf{Non-defining relative clauses:}

On the contrary, non-defining relative clauses do not filter anything out. It is just a convenient
place to add some additional extra information to some already well-defined/well-known item.

\begin{ItemizeWithOrigParIndent}
   \item Use non-defining relative clause to add extra non-essential information.
   \item Use a comma before the non-defining relative clause. Use a comma or a full-stop after it.
   \item Spell it with pause in speech.
\end{ItemizeWithOrigParIndent}

Here is examples:
\begin{ItemizeWithOrigParIndent}
   \item ``The children, who love films, went to the cinema.''~--- all of the children went
   to the cinema. Дети (а они любят фильмы) пошли в кино.

   ``The children who love films went to the cinema.''~--- only some of the children went
   to the cinema. Дети, (только те) которые любят фильмы, пошли в кино.

   \item ``Her brother, who is a musician, lives in New York.''~---she has one brother.
   He lives in New York. Её брат (а он музыкант) живёт в Нью-Йорке.

   ``Her brother who is a musician lives in New York.''~--- she has more than one brother.
   One of them lives in New York. Её брат, тот который музыкант, живёт в Нью-Йорке.

   \item ``My house, which is by the beach, has a great view.''~--- I have only one house.
   Мой дом (а он находится у пляжа) имеет прекрасный вид.

   ``My house which is by the beach has a great view.''~--- I have more than one house.
   Мой дом, тот из них который у пляжа, имеет прекрасный вид.

   \item ``They went to the only school in the village, which had good teachers.''~--- there
   was only one school in the village. Они ходили в единственную школу в деревне
   (и в ней были хорошие учителя).

   ``They went to the only school in the village which had good teachers.''~--- there was
   more than one school in the village, but only one had good teachers. Они ходили в единственную
   школу в деревне, (ту из них) которая имела хороших учителей.
\end{ItemizeWithOrigParIndent}


\subsubsection{Vocabulary: Neighbours}
\begin{description}[leftmargin=4.1cm,style=nextline,before={\renewcommand\makelabel[1]{##1~---}}]
  \item[\ex{mind my own business}] refrain from meddling (вмешательства) in other people's affairs
  \item[\ex{keep myself to myself}] to spend time alone, not talking to other people very much
  \item[\ex{invite somebody over}] invite my neighbour over for a coffee~--- пригласить к себе
  \item[\ex{make friends with}] we made friends with our \\neighbours immediately
  \item[\ex{get on my nerves}] действовать на нервы
\end{description}
\vspace{-\parskip}
\begin{description}[leftmargin=2.2cm,style=nextline,before={\renewcommand\makelabel[1]{##1~---}}]
  \item[\ex{chant}] ритуальное песнопение, монотонно говорить
  \item[\ex{nuisance}] надоедливый человек, зануда, досада
  \item[\ex{nosy}] любопытный
  \item[\ex{cloak}] плащ, мантия
\end{description}

\subsubsection{Vocabulary: Make Yourself at Home}
\begin{description}[leftmargin=4.7cm,style=nextline,before={\renewcommand\makelabel[1]{##1~---}}]
  \item[\ex{Make yourself at home}] you want someont to feel relaxed in your house
  \item[\ex{Put your feet up}] you know someone is tired and you want them to relax
  \item[\ex{Excuse the mess}] you are apologising for how your home looks
\end{description}
\vspace{-\parskip}
\begin{description}[leftmargin=3.0cm,style=nextline,before={\renewcommand\makelabel[1]{##1~---}}]
  \item[\ex{Be my guest}] you give permission for someone to use something that is yours
  \item[\ex{Help yourself}] you are offering someone food
  \item[\ex{Have a seat}] you inviting someone to sit down
\end{description}







\end{multicols}











%================================================================================================
%================================================================================================
%================================================================================================

\begin{multicols}{2}[\section{SpeakOut Upper-Intermediate}]

\end{multicols}













%================================================================================================
%================================================================================================
%================================================================================================

\begin{multicols}{2}[\section{SpeakOut Advanced}]

\end{multicols}










%================================================================================================
%================================================================================================
%================================================================================================

\setcounter{secnumdepth}{2} % Enumerate only Sections (1) and SubSecions (2)
\begin{multicols}{2}[\section{Phrases from speech}]

%\subsection{Handy Fixed Expressions}
%\begin{ItemizeWithOrigParIndent}
%\end{ItemizeWithOrigParIndent}
%\subsection{Vocabulary}
%Examples with incorrectly Used or forgotten vocabulary or complicated cases.
%\begin{ItemizeWithOrigParIndent}
%\end{ItemizeWithOrigParIndent}
%\subsection{Preposition}
%\begin{ItemizeWithOrigParIndent}
%\end{ItemizeWithOrigParIndent}
%\subsection{Grammar}
%\subsection{Long phrases}
%\begin{ItemizeWithOrigParIndent}
%\end{ItemizeWithOrigParIndent}


\subsection{Handy Fixed Expressions}
\begin{ItemizeWithOrigParIndent}
   \item He can buy whatever \ex{his heart desires}.
   \item ``I have broken your vase''

   ``\ex{It is nothing}''~--- it is Ok (ничего страшного).

   \item ``Thank you for the dinner!''

   ``\ex{Do not mention it}''~--- не за что.

\end{ItemizeWithOrigParIndent}

\subsection{Vocabulary}
Examples with incorrectly Used or forgotten vocabulary or complicated cases.

\begin{ItemizeWithOrigParIndent}
   \item Trains in tube share the same \sout{railway} \ex{track}.
   \item I live on the second \ex{floor}. But: My house has four \ex{storeys}.
   \item I wish they set more \ex{lanterns} (фонари) on the streets.
   \item \ex{be engaged}. \ex{get engaged into} an activity (into drug dealing).
   \item По большей части~--- \ex{for the most part} / \ex{mostly}.
\end{ItemizeWithOrigParIndent}


\subsection{Preposition}
\begin{ItemizeWithOrigParIndent}
   \item There was a huge \ex{demand on} the features YouTube was providing.
   \item Do not \ex{waste time on} V\textsubscript{ing} (doing) / noun (laundry).
\end{ItemizeWithOrigParIndent}


\subsection{Grammar}
\subsubsection{He suggested that he go/goes}
   He suggested that I / you / we / \ex{he} / \ex{she} / \ex{it} \sout{does} \ex{do}. In other words in phrases like \ex{prefer that, suggest that, vote that, wish that} and so on\dots\ after 3rd person pronouns (he/she/it) you are to use verbs withoud -s/-es endings.

   Here is another, more complete explanation.

   \textit{She suggested that he go to the cinema.} and \textit{She suggested that he goes to the cinema.} are both correct, but they have different meanings.

   Here's how she might suggest that \ex{he goes} to the cinema:

   ALICE: ``Where do you think he goes every Thursday evening?''

   JANE: ``Hmm\dots\ well\dots\ cinema tickets are cheap on Thursdays, and he loves movies. Maybe he goes to the cinema?''

   Here's how she might suggest that \ex{he go} to the cinema:

   BOB: ``I want to go out and have some fun this evening. Do you have any suggestions?''

   JANE: ``Why don't you go to the cinema?''


\subsection{Long phrases}
\begin{ItemizeWithOrigParIndent}
   \item Паспорт на то и паспорт, что бы в нём ставили отметки. That's what passports are
   for~--- getting stamped.

   \item Проблема заключается не в том, что фирма одна и та же, а в том, что в том календаре~--- петухи.

   The mistake is not that manufacturer of the calendars is the same, but that there are
   roosters in later one.

   \dots\ заключается не в том что\dots, а в том что\dots:
   \begin{ItemizeWithOrigParIndent}
      \item \dots\ (The reason why you feel) \ex{is not because}\dots\ (you're outside gravity),
      \ex{but because}\dots\ (you're no longer resisting)

      \item \dots\ (The advantage of this representation) \ex{is not that}\dots\ (it gives us),
      \ex{but that}\dots\ (it enables us)

      \item \dots\ (Her purpose) \ex{was not to}\dots\ (level accusations),
      \ex{but to}\dots\ (promote human rights)
   \end{ItemizeWithOrigParIndent}


\end{ItemizeWithOrigParIndent}
\end{multicols}
















%================================================================================================
%================================================================================================
%================================================================================================

\newpage
\setcounter{secnumdepth}{3} % Enumerate only Sections (1) and SubSecions (2)

\begin{multicols}{2}[\section{The United Kingdom. 2016.}]

\subsection{Phrases from speech}

\subsubsection{Handy Fixed Expressions}
\begin{ItemizeWithOrigParIndent}
   \item That would work as well.
   \item Take the second right. Excuse me, please, could you (give me directions) /
   (tell me the way) to The London Eye?
   \item Ask something in caffee: Please, could I have a green tea? I would like a green tea, please.
\end{ItemizeWithOrigParIndent}


\subsubsection{Vocabulary}
Examples with incorrectly Used or forgotten vocabulary or complicated cases.
\begin{ItemizeWithOrigParIndent}
   \item Do many people here \sout{ride} \ex{commute/travel} to London?
   \item We drink \sout{much} \ex{a lot of} tea.
   \item \sout{Two times in a week} \ex{Twice} a week.
   \item \sout{Previous year.} \ex{Last} year.
   \item We can show you one \sout{our} \ex{of ours} videos.
   \item You \sout{may not} \ex{doesn't (need/have) to} search.
   \item I'll make two \sout{circles} \ex{circuits} around the church.
   \item Zebra is an animal, that has black and white \sout{lanes} \ex{stripes} on its body.
   \item As soon as we \sout{entered} \ex{got on} the train.
   \item Which one? \sout{Second one.} \ex{Later (formerly).}
   \item Finally, we \sout{exited on} \ex{traveled to} Westminster station.
   \item Сделать мэйкап: \sout{Make make-up.} \ex{Do make-up. / Put on make-up.}
   \item There are two \sout{parts} \ex{groups} of people in Korolev.
   \item Подстричь газон: \sout{Cut the grass.} \ex{Mow the lawn}.
   \item She wears a coat to \sout{protect herself from cold} \ex{keep warm}.
   \item Exam: Pass an exam. Fail an exam. Do well/badly in an exam.
\end{ItemizeWithOrigParIndent}

\subsubsection{Preposition}
\begin{ItemizeWithOrigParIndent}
   \item \sout{Ask help. Ask to help.} \ex{Ask for help}.
\end{ItemizeWithOrigParIndent}

\subsubsection{Grammar}
\begin{ItemizeWithOrigParIndent}
   \item \sout{Have he married?} \ex{Did he get married?}
\end{ItemizeWithOrigParIndent}

\subsubsection{Long phrases}
\begin{ItemizeWithOrigParIndent}
   \item We were impressed \sout{that there was a very quiet train} by how quiet the train was.
   \item Не могли бы вы поставить (тарелку, например. За столом): Could you please, put it down.
   \item Он подсадил меня на YouTube: He has got me addicted to YouTube.
   \item Хочешь я куплю тебе что нибудь, что бы ты повеселела? Do you want me to buy something
   for you (to make you feel better?) / (that helps you?)
\end{ItemizeWithOrigParIndent}



%================================================================================================
\subsection{Pronunciation}
\begin{ItemizeWithOrigParIndent}
  \item \textbf{-ng}: ki\textbf{ng}, swimmi\textbf{ng}, includi\textbf{ng}.
  \item s\textbf{i}x: и.
  \item \textbf{h}im, \textbf{h}er, \textbf{h}is, \textbf{h}ave.
  \item \textbf{w}ere: \sout{в}[еэ]. \textbf{у[вф]}[еэ].
\end{ItemizeWithOrigParIndent}



%================================================================================================
\subsection{Vocabulary}

\subsubsection{New words, phrases}
\begin{description}[leftmargin=\OriginalParIndent,rightmargin=1mm, style=nextline, before={\renewcommand\makelabel[1]{##1~---}}]
  \item[Allotment] брит. небольшой участок земли, сдаваемый в аренду под огород.
  \item[Kettle] чайник.
  \item[Teabag] чайный пакетик.
  \item[Hedge] живая изгородь.
  \item[Hen] курица.
  \item[Gravy] подливка (из сока жаркого), соус, сок.
  \item[Admit one] ticket for one.
  \item[Stroll around] wander.
  \item[Afterward, beforehand] не надо ставить после них существительного (в отличие от after, before),
                             например: \textit{What will you do afterwards?}
  \item[Riddle] загадка, тайна.
  \item[Tame (squirrels)] приручённый, одомашненный.
  \item[Wages] зарплата.
  \item[Renovate] обновлять. \textit{Обновленная кухня}, \textit{реконструированный парк}~---
  \textit{renovated kitchen}, \textit{renovated park}.
  \item[Do a chore] делать повседневную работу по дому.

  \item[Flat] квартира в многоквартирном доме.
  \item[Apartment building] многоквартирные дом.
  \item[House] частный дом.

  \item[Go/come] TODO.
  \item[Take/bring] TODO.
\end{description}

\subsubsection{Phrases verbs}
\begin{description}[leftmargin=2.2cm,style=nextline,before={\renewcommand\makelabel[1]{##1~---}}]
  \item[Call off] cancel (a meeting).
  \item[Use up] израсходовать.
  \item[Clean out] ``обчистить''.
  \item[Put on] надеть.
  \item[Take off] раздеться.
  \item[Make up] invent, выдумать.
  \item[Leave out] skip, пропускать, не принимать во внимание.
  \item[Lock in] запереть (себя внутри).
  \item[Bring up] поднимать вопрос (неприятный).
  \item[Hand out] раздавать.
  \item[Hand in] передавать (в руки).
  \item[Turn over] переворачивать.
  \item[Jot down] быстро сдампить мысли на бумагу.
\end{description}

\subsubsection{Saying, idioms}
\begin{ItemizeWithOrigParIndent}
  \item \ex{Bad hair day}~--- a day on which everything seems to go wrong.
  \item \ex{Can't stand / Can't bear}~--- means to be unable to tolerate somebody or something.
  \item \ex{Watched pot never boils}~---something we wait for with impatient attention seems to take forever.
  \item \ex{Jump on the bandwagon}~--- if you jump on the bandwagon, you join a growing movement in support
  of someone or something when that movement is seen to be about to become successful.
  \item \ex{The pot calling the kettle black}~---an idiom used to claim that a person is guilty of the very thing of which they accuse another.
\end{ItemizeWithOrigParIndent}

\subsubsection{Degree of likability}
\begin{itemize}
  \item Despise~--- презирать.
  \item Abhor/Detest~--- ненавидеть, питать отвращение; не выносить, не терпеть.
  \item Hate.
  \item Can't stand / Can't bear~--- means to be unable to tolerate somebody or something.
  \item Dislike/don't like.
  \item Be afraid of
  \item To be easy: \textit{I am easy} (informal).
  \item Do not mind~--- it's ok. To \textit{mind} something means, that you are distressed, annoyed or worried
  by it. Examples:
  \begin{itemize}
    \item \textit{Do you mind V\textsubscript{ing}? No, I don't mind.}
    \item \textit{Do you mind helping me with my homework? I don't mind helping you with\dots}
  \end{itemize}
  \item Quite like / quite enjoy.
  \item Like/enjoy.
  \item Love.
  \item Adore.
  \item Worship~--- боготворить, поклоняться.
\end{itemize}
\vspace{-\parskip}
It seems that almost all of these words must be followed by gerund or noun, see \ref{sed:Verbs_followed_by_gerunds}.


%================================================================================================
\subsection{Riddle}
\textit{Smith, where Jones had had ``had'', had had ``had had''. ``Had had'' had had examiner's approval.}




%================================================================================================
\subsection{Pronouns}
\textit{I don’t like vegetables, I don't like \sout{it} \textbf{them}.}



%================================================================================================
\subsection{Plural forms}
Nouns ending in -o: \textit{tomato\ex{es}}.\\
Irregular: \ex{foot~-- feet}, \ex{tooth~-- teeth}.\\
Does not change in plural: \ex{deer}, \ex{shrimp}, \ex{moose}, \ex{fish}.




%================================================================================================
\subsection{Expressing probability}
Might, could.\\
\ex{Might} not necessarily means past! \ex{Could} implies several choices: \textit{We could go to the cinema, or we could listen to music.}






%================================================================================================
\subsection{Articles}

\subsubsection{Fixed expressions}

\begin{minipage}{0.32\linewidth}
the sea\\
the environment\\
the country\\
the Internet\\
the cinema\\
the theatre\\
the opera\\
the radio\\
the navy\\
the army\\
the air force\\
the police
\end{minipage}
\begin{minipage}{0.65\linewidth}
the bank\\
the post office\\
the cinema\\
the station\\
the store\\
the doctor\\
the dentist\\
the airport: \textit{they are building a new airport} (a kind of airport)\textit{, because the
old airport is not big enough} (a specific one)
\end{minipage}

\begin{multicols}{3}
at the top\\
at the bottom\\
in the middle\\
at the centre\\
on the right/left\\
at the end
\end{multicols}

\begin{multicols}{3}
\za lunch\\
\za breakfast\\
\za dinner
\end{multicols}

\textit{He will go at \za school.} Если можно переформулировать без school, например: он пойдет учиться.\\
\textit{He will go at \textbf{the} school, where his parents work.} Если имеется в виду конкретное здание (specific place), а не процесс.\\
Вместо \ex{school} могут быть: \ex{hospital}, \ex{bed}, \ex{university}, \ex{jail}, \ex{prison}, \ex{college}.



\subsubsection{Buildings / Places.}
Places used \textbf{without the}:
\begin{description}[leftmargin=2.9cm,style=nextline,before={\renewcommand\makelabel[1]{##1:}}]
  \item[Streets, roads] \za Fleet Street.
  \item[Airports] \za Heathrow airport.
  \item[Parks] \za Hyde Park.
  \item[Squares] \za Times Square.
\end{description}
\vspace{-\parskip}
Name + building: \textit{\za Oxford University}, \textit{\za San Diego Zoo}, but \textit{\ex{The} Royal Academy} and \textit{\ex{The} Provincial Courts} because \textit{Royal} and \textit{Provincial} are adjectives.




\vspace{2mm}
Places used \textbf{with the}:
\begin{description}[leftmargin=3.9cm,rightmargin=\OriginalParIndent,style=nextline,before={\renewcommand\makelabel[1]{##1:}}]
  \item[Hotels] The Hilton Hotel.
  \item[Theaters, cinemas, operas] The Sydney Opera.
  \item[Museums] The British Museum.
  \item[Monuments, towers] The London Eye.
  \item[Names with ``of''] The University of Oxford. (But, Oxford University, see above.)
\end{description}
\vspace{-\parskip}
If any of those (restaurant, hotel, \dots) named after people, whose name ends with \ex{-s}, \ex{-s’} or \ex{-’s} do not use \ex{the}: \ex{\za St James’ Church}.



\subsubsection{Companies / Organizations.}
We do not use \ex{the} with names of companies:\\
\textit{\za Google}, \textit{\za Adidas}.

We often use \ex{the} with names of organizations \& newspapers:
\textit{\ex{The} United Nations}, \textit{\ex{The} New York Times}.



\subsubsection{Geographical areas.}
We \textbf{do not use articles with}:

\hspace{\OriginalParIndent}\begin{minipage}{0.55\linewidth}
Cities, towns, villages.\\
Countries, provinces, states.\\
Continents.
\end{minipage}
\begin{minipage}{0.3\linewidth}
Islands.\\
Mountains.
\end{minipage}\vspace{\parskip}


We \textbf{use \ex{the} with}:

\hspace{\OriginalParIndent}\begin{minipage}{0.55\linewidth}
Oceans, seas.\\
Rivers, canals, straits.\\
Deserts.
\end{minipage}
\begin{minipage}{0.3\linewidth}
Groups of islands.\\
Mountain ranges.
\end{minipage}\vspace{\parskip}

We also \textbf{use \ex{the}} with compass directions: \textit{The North}, \textit{The North-East}, \textit{The East}, \dots: \textit{\ex{The} North of Germany}, \textit{\ex{The} East of Africa}, \dots\\
Do not mix up situations in which compass direction is part of name:
\textit{\za South America} (name of continent), \dots\\
But \textbf{do not use \ex{the}} with \textit{Southern}, \textit{Northern}, \dots







%================================================================================================
\subsection{Prepositions}

\begin{ItemizeWithOrigParIndent}
  \item Mr Jones is not here now. He is \ex{at} lunch.
  \item Tony is waiting for you \ex{at} the restaurant.
  \item The speep are \ex{in} the field.
  \item Hens are \ex{in} the garden.
\end{ItemizeWithOrigParIndent}

\begin{ItemizeWithOrigParIndent}
  \item \ex{On} train.
  \item \ex{On} the underground.
  \item \ex{By} bus.
  \item \ex{On} foot.
\end{ItemizeWithOrigParIndent}






%================================================================================================
\subsection{That/this/those/these.}
\textit{\ex{That} milk in the fridge is more than two weeks old.} \ex{That} is used because milk in
the fridge~--- fridge is closed.

\subsubsection{That's ok}
\begin{ItemizeWithOrigParIndent}
  \item Sorry about \ex{that}.
  \item \ex{That} was a terrible film. (But \textit{\ex{This} is a terrible film}~---
  \textit{is}~--- we are watching it now).
  \item I didn't know \ex{that}.
  \item \ex{That's} wonderful.
\end{ItemizeWithOrigParIndent}





%================================================================================================
\subsection{Linking words}
\subsubsection{Besides}
As a preposition \ex{besides} has the same meaning as ``as well as'' and ``in addition to'' .\\
\textit{\ex{Besides} being a magnificent chief, he also paint extremely well.}\\
\textit{\ex{Besides} the clarinet, she also plays the flute.}

Do not mix:\\
Beside~--- next to, near, close to.\\
Besides~--- in addition to.

Let's pretend you had a long day at work, and you are tired. You just want to stay at home and watch TV. Your friend calls and asks ``do you want to join me for a dinner?''.
Your answer: \textit{No, thanks. Today was a long day and I'm tired. \ex{Besides}, I just ate dinner.}





%================================================================================================
\subsection{Relative clauses}

Correction bad/incorrect relative clauses:
\begin{ItemizeWithOrigParIndent}
  \item The tree \sout{whose} leaves have fallen off, is really old --> The tree \textbf{that has lost} all its
  leaves, is really old.
  \item I prefer to eat at restaurant \sout{whose} tables aren't very close together. --> I prefer to eat at restaurant \textbf{with tables, that} are further apart. / I prefer to eat at restaurant \textbf{where the tables} are further apart.
  \item That is the door, \sout{whose} lock doesn't work. --> That is the door \textbf{with lock, that} doesn't work.
\end{ItemizeWithOrigParIndent}





%================================================================================================
\subsection{Verbs followed by gerunds only.} \label{sed:Verbs_followed_by_gerunds}
These verbs are commonly followed by gerunds:
\vspace{-4\parskip}
\begin{multicols}{4}
\begin{itemize}[leftmargin=3.05mm]
  \item postpone
  \item delay
  \item finish
  \item feel like
  \item enjoy
  \item tolerate
  \item imagine
  \item can't help
  \item involve
  \item avoid
  \item keep
  \item anticipate
  \item complete
  \item admit
  \item report
  \item defend
  \item mention
  \item deny
  \item suggest
  \item discuss
  \item recall
  \item consider
  \item recommend
  \item risk
  \item resist
  \item mind
  \item practice
  \item despise
  \item dislike
  \item appreciate
  \item quit
\end{itemize}
\end{multicols}
\vspace{-5\parskip}
\textit{Doctor \ex{recommended} \sout{you} \ex{eating} healthy food.}\\
\ex{Feel like} is always followed by a gerund.\\
To \textit{\ex{feel like} doing something} means to be in the mood to do something.\\
\ex{can't help doing}~--- не мочь удержаться от чего-либо.\\
\textit{He \ex{admitted} stealing money from her pursue}~--- Он признался в краже денег из её кошелька.








%================================================================================================
\subsection{Wishes.}

I wish I\dots
\begin{tabular}{l@{\hspace{5.5mm}} l@{\hspace{4mm}} l}
\hline
future  & \textit{would + inf}   & \multirow{2}{*}{\textit{could + inf}}        \\
present & past simple   &                                     \\
\hline
past    & past perfect  & \textit{could + have + V\textsubscript{3}}   \\
\hline
\end{tabular}

\textbf{Notes:}

\textit{could + have + V\textsubscript{3}}~--- often means disappointment (definitely impossible to fix): \textit{I wish I could have gone to the party}.

\textit{wish \dots\ wouldn't \dots}~--- for complains about something that happens over and over again: \textit{I wish out neighbours wouldn't play that terrible music}.

\textbf{Be careful with future!}\\
It is not actually future.
\textit{would} in table above is \textit{past(will)}, where ''will`` means ``show willingness''. That's why following sentences are incorrect:
\begin{ItemizeWithOrigParIndent}
  \item \textit{\sout{I wish I would know her.}} Дословный перевод: \textit{Я желаю что бы я зажелал её узнать}.
  \item \textit{\sout{I wish I would have a new phone.}} Дословный перевод: \textit{Я желаю что бы я зажелал заиметь новый телефон}.
\end{ItemizeWithOrigParIndent}
\dots\ and should be fixed as follows:
\begin{ItemizeWithOrigParIndent}
  \item \textit{I wish I knew her.}
  \item \textit{I wish I had a new phone.}
\end{ItemizeWithOrigParIndent}

Where ``will'' means a future event, we cannot use ``wish'' and must use ``hope'':
\begin{ItemizeWithOrigParIndent}
  \item \textit{There's a strike tomorrow. I hope some buses will still be running.}
  \item \textit{I hope everything will be fine in your new job.}
\end{ItemizeWithOrigParIndent}






%================================================================================================
\subsection{Probability \& Modal Verbs}

\begin{tabular}{l@{\hspace{4mm}} l@{} l@{} l@{} l@{}}
tense   & rule \\
\hline
present & subj + one of modal verbs & \ +      & \ be     & \ + V\textsubscript{ing} \\
past    & subj + one of modal verbs & \ + have & \ been & \ + V\textsubscript{ing}   \\
\hline
\end{tabular}

Below is a list of modal verbs:
\vspace{-7\parskip}
\begin{multicols}{3}
\begin{ItemizeWithOrigParIndent}
  \item must
  \item may
  \item might
  \item could
  \item should
  \item ought to
\end{ItemizeWithOrigParIndent}
\end{multicols}




\end{multicols}

\end{document}
